---
title: "Air passengers 1949-1960"
output: pdf_document
geometry: margin=.75in
graphics: yes
header-includes:
- \usepackage{graphicx}
- \usepackage{color}
fontsize: 9pt
classoption: letter

---

**22 marks**

**Airline passengers**.   
Here we will examine the monthly total of the number of air passengers in the US from 1949 to 1960.

The data are available as the data set `AirPassengers` from the `datasets` package in the standard `R` distribution.

```{r setup data}
AirPassengers
```

a. **(7 marks)** Use the function `decompose()` on this data and plot the results.  

    i. *(2 marks)* Show your plot.
    
    ```{r}
    plot(decompose(AirPassengers))
    ```
    
    ii. *(1 mark)* Describe the trend.
    
        **There is an rising trend with the time increases.**
    
    iii. *(2 marks)* Describe the seasonal pattern.
    
    
         **The seasonal is in a repeating cycle across years. Graphs are indentical every
        year.**
    
    iv. *(2 marks)* What do you conclude from the residuals?
        
        **We don't see a clear pattern in residuals. However, the plot between 1954 and 1958 is different from other years.**
    

b.  **(9 marks)** Use the function `stl()` on this data with seasonality loess span `s.window = 7` and the local polynomial for the seasonal loess being a line.

    i. *(2 marks)* Show your plot.
    
    ```{r}
    plot(stl(AirPassengers, s.window=7, s.degree=1))
    ```
    
    
    ii. *(2 marks)* Describe the trend.  Is it significant? Why or why not?
    
        **It's still an rising trend. However, it is significant because the variation is really large as the size of the range rectangle is smaller than the remainder's.**
    
    iii. *(3 marks)* Describe the seasonal pattern.  Is it significant? Why or why not?
    
         **Seasonal variation gets larger and larger but still in certain cycles. It is significant because comparing to remainder, it also has small size of range rectangle.**
        
    iv. *(2 marks)* What do you conclude from the residuals?
        
        **We can't see a clear pattern in residuals. There are variations across years.**

c. **(6 marks)** Compare the plots from parts (a) and (b).

    i. *(2 marks)* What are the major differences in the plots?
       
       **Seasonal variation in (b) increases across years comparing to constant variation in (a).**
       
    ii. *(2 marks)* What characteristics of the two methods caused these differences?
    
        **It is because for (a), we are using simple average for months in all years. In (b), we are smoothing the value in a 7-year window and use the smooth values as estimate.** 

    iii. *(2 marks)*  Which fitted model do you prefer and **why**?
    
         **I prefer the second one because we are able to find a descriptive pattern in seasonal. Residuals in the second model is random, which is a desired result.**
        